Errata

### 2025-04-07 tex/Manuscript-4.tex
Refactor code structure for improved readability and maintainability

\begin{center}
  \textbf{\Large Manuscript 4}
\end{center}  
\vspace{0.1cm}
\begin{center}
  \textbf{\huge Microbially induced soil aggregate turnover across different climates and moisture regimes}

    \vspace{0.2cm}
    \textit{Geoderma}\\
    \textit{DOI 10.2139/ssrn.4740477}

@@

  \end{center}
  \begin{justify}
    Julia Mitzscherling\textsuperscript{1}, Rómulo Oses\textsuperscript{2}, Nicolás Riveras-Muñoz\textsuperscript{3}, Carsten W. Müller\textsuperscript{4,5}, Oscar Seguel\textsuperscript{6},   Peter Kühn\textsuperscript{3}, Thomas Scholten\textsuperscript{3} and Dirk Wagner\textsuperscript{1,7}
  \end{justify}

@@

  \vspace{0.4cm}
  \begin{flushleft}
    Received: 27 February 2024
  \end{flushleft}
  \cleardoublepage

@@

In the frame of the EarthShape project (German Science Foundation SPP 1803), soils were studied at the following four study sites distributed over \SI{1300}{\kilo\metre} along the Chilean Coastal Cordillera: Pan de Az{\'u}car National Park (\(\sim\)\ang{26}S), Santa Gracia Nature Reserve (\(\sim\)\ang{30}S), La Campana National Park (\(\sim\)\ang{33}S), and Nahuelbuta National Park (\(\sim\)\ang{38}S). 
The study sites cover a climate gradient with decreasing annual temperature and increasing annual precipitation from north to south and corresponded to arid, semi-arid, mediterranean and humid climates (\cref{fig:figure_1_label}) resulting in different soil moisture regimes \citep{Bernhard2018}. 
Due to the long-term impact of stable climate conditions \citep{Ewing2006} soils show different stages of soil development. 
They were classified into Regosols in Pan de Az{\'u}car, Cambisols in Santa Gracia and La Campana, and Umbrisols in Nahuelbuta \citep{Bernhard2018b}. 
Vegetation types range from very sparse higher plant desert vegetation in the north, where biocrust soil coverage can reach as much as \SI{40}{\percent} of the area (biocrust-dominated desert vegetation) over semi-desert open shrubland and mediterranean dry forest and to temperate mixed broadleaved-coniferous forest in the south \citep{Bernhard2018a, Bernhard2018b}.

Surface depth increments (\SIrange{0}{5}{\centi\metre}) were collected in March--April 2016 on the mid slope position of the South-facing hillslopes in each study site and were directly sieved (\SI{< 2}{\milli\metre}) on ethanol-sterilized sieves to remove coarse organic material and stones. 
Samples were stored at \SI{4}{\degreeCelsius} until further processing. 
Detailed site-specific descriptions and soil classifications were performed by \citet{Bernhard2018a}, \citet{Bernhard2018b} and \citet{Oeser2018a}, \citet{Oeser2018b}.

@@

Soil-chemical properties were analyzed for the native and the sterilized soils. 
The total organic carbon (TOC) was determined by Potsdamer Wasser- und Umweltlabor GmbH (PWU, Potsdam, Germany) according to DIN EN ISO/IEC 17025:2005. 
The aggregate size separation was done by wet sieving adapted from \citet{Elliott1986} as described in \citet{RiverasMunoz2022}. 
In the first step, \SI{10}{\gram} of soil material was irrigated with \SI{50}{\milli\liter} water. 
A series of four sieves, which were placed in descending order, was used to obtain the following soil aggregate size classes: 
(i) \SI{> 500}{\micro\metre} (large macroaggregates); 
(ii) \SIrange{250}{500}{\micro\metre} (small macroaggregates); 
(iii) \SIrange{53}{250}{\micro\metre} (large microaggregates); 
(iv) \SIrange{20}{53}{\micro\metre} (small microaggregates), and 
(v) \SI{< 20}{\micro\metre} (silt/clay). 
Aggregates were separated manually by moving the sieve up and down 30 times during \SI{5}{\minute} in a \SI{2.5}{\liter} water tank. 
The aggregate classes retained on each sieve were oven-dried at \SI{105}{\degreeCelsius}, and weighed. 
In the following, large and small macroaggregates are referred to as macroaggregates (MAC), large and small microaggregates are referred to as microaggregates (MIC) and the silt/clay class is referred to as SC.

@@

Sequencing raw data are publicly available via the European Nucleotide Archive under the Project accession number \texttt{PRJEB72146}: \url{https://www.ebi.ac.uk/ena/browser/view/PRJEB72146}. 
All data used in this study are freely available under the Creative Commons Attribution 4.0 International (CC BY 4.0) open access license at GFZ data services. 
All samples are identified via a unique IGSN (international geo sample number). 
These data are available as a supplementary dataset in \citet{Bernhard2018b} with the following DOI: \url{https://doi.org/10.5880/GFZ.5.3.2018.001}. 
When using the data please cite: \citet{Bernhard2018a}.

@@

Macroaggregates (MAC), microaggregates (MIC) and the silt/clay (SC) classes were differently distributed in arid, semi-arid, mediterranean and humid soils (\cref{fig:figure_2a_label}). 
Less developed soils from arid and semi-arid climate conditions were characterized by the lowest content of MAC ranging around \SIrange{60}{65}{\percent}. 
Semi-arid soils displayed a higher MIC content and a lower SC content compared to arid soils. 
More developed soils from the mediterranean and humid climates were characterized by the highest abundance of MAC with nearly \SI{80}{\percent} in mediterranean soils and more than \SI{90}{\percent} in humid soils. 
Accordingly, the proportion of MIC was low in mediterranean soils (\(\sim\)\SI{20}{\percent}) and lowest in humid soils (\SI{< 20}{\percent}). 
Moreover, SC was hardly present in mediterranean and humid soils (\(\sim\)\SI{2}{\percent}).

@@

The SOC and TN (\cref{fig:figure_2b-c_label}) contents were lowest in the less developed soils and increased with soil development along the climate gradient, with the lowest SOC (\(\sim\)\SI{1}{\percent}) and TN values (\(\sim\)\SI{0.1}{\percent}) in arid soils and highest values in humid soils (\(\sim\)\SI{10}{\percent} and \(\sim\)\SI{0.5}{\percent}). 
In each soil, the SOC and TN contents were lowest within the MAC class and highest within the SC class. 
The C/N ratios (\cref{fig:figure_2d_label}) were highest in arid soils. 
Lowest values were observed in semi-arid soils with an increasing trend over mediterranean to humid soils. 
Within each soil, the highest C/N ratios can be observed within MAC, followed by SC and MIC.

The aggregate stability (MWD) was lowest in arid and semi-arid soils, increased in mediterranean soils and peaked in humid soils (\cref{fig:figure_a1_label}).

A principal component analysis (PCA) including soil aggregate size classes, aggregate stability (MWD), SOC, TOC and TN content as well as the C/N determined during repeated W-D cycles (\cref{fig:figure_3a_label}) revealed similar soil properties in native and sterilized soils from the same climatic condition. 
The separation of soils along the x-axis, explaining \SI{66.0}{\percent} of the variance, was in accordance with the climate gradient from arid to humid conditions. 
The difference between arid and semi-arid soils was mainly visible on the y-axis, explaining \SI{27.6}{\percent} of the variance. 
In addition, native and sterilized soils from arid conditions were separated along the y-axis.

@@

Differences between native and sterilized soils were further evaluated in separate PCA ordinations, showing significant differences between native soils and sterile controls not only in arid (\(p = 0.005\)), but also in semi-arid (\(p = 0.008\)) and mediterranean soils (\(p = 0.02\)) (\cref{fig:figure_4_label}, \cref{tab:table_a1_label}). 
At these three sites, the native samples with microorganisms before W-D treatment (\(\square\)R0) were similar to all sterile controls (\(\circ\)R0--R6, CM) and were thus referred to as microbially not (yet) influenced. 
After at least 1 W-D cycle, microbial samples clustered apart from the sterile controls and were thus claimed to be microbially influenced. 
The difference between microbially influenced and not influenced samples appeared mainly along the x-axis of the plots, which explained the highest variance. 
In arid soils, the microbially influenced soils were characterized by higher C/N ratios in all aggregate classes (\cref{fig:figure_4a_label}). 
In semi-arid and mediterranean soils, soil aggregate sizes in microbially influenced soils were characterized by higher levels of MIC and C/N\textsubscript{SC}, and lower levels of MAC compared to the sterile controls (\cref{fig:figure_4b-c_label}). 
Microbially influenced humid soils were not significantly different from sterile soils. 
However, the majority of microbially influenced soils was characterized by higher MAC and MWD, while sterile soils had higher MIC contents (\cref{fig:figure_4d_label}).

@@

Based on 16S and 28S rRNA gene copy numbers, bacteria were most abundant among all four climate conditions, followed by archaea and fungi (\cref{fig:figure_5a_label}). 
Only in humid soils, fungi were slightly more abundant than archaea and possessed highest abundances among soils. 
The lowest abundances of bacteria, archaea and fungi were observed in arid soils. 
Still, they increased over the course of the experiment with a peak after 3 cycles and maximum values under constant moisture. 
Semi-arid and mediterranean soils were characterized by highest abundances that remained relatively constant during the W-D cycles and constant moisture. 
In humid soils, bacterial and archaeal abundances were stable and fungal gene copies increased after 3 W-D cycles. 
However, all gene abundances of humid soils were lowest in soils exposed to constant moisture.

Linear (Pearson) and non-linear (Spearman) correlation analyses (\cref{fig:figure_a2_label}) revealed that bacterial and archaeal abundances were positively correlated with each other. 
Both were inversely correlated with the C/N ratios in each soil class. 
On the contrary, the fungal abundance was positively correlated with MAC and MWD, while being inversely correlated with MIC and the S/C class. 
Furthermore, fungi were positively correlated with the SOC and TN contents in each aggregate class.

@@

Similar to their abundance, bacterial and archaeal diversity based on the alpha-diversity indices Chao-1 and Shannon (\(H\)) was highest in semi-arid soils from Santa Gracia (median \(H = 7.4\)) and in mediterranean soils of La Campana (\(H = 7.5\)) with maximum diversity indices in mediterranean soils (\cref{fig:figure_5b_label}). 
Although lowest microbial abundances were observed in arid soils, the bacterial and archaeal diversity was lowest in humid soils from Nahuelbuta (\(H = 6.0\)), followed by arid soils from Pan de Az{\'u}car (\(H = 6.6\)). 
In arid and semi-arid soils, the diversity according to Chao-1 increased after one W-D cycle and decreased afterwards. 
Contrary, mediterranean and humid soils were characterized by an increasing diversity during W-D cycles, but possessed a low diversity in the sample exposed to constant moisture. 
Like the bacterial and archaeal abundance, the number of taxa and the diversity were negatively correlated with the C/N ratios in all soil classes (\cref{fig:figure_a3_label}).

@@

Like the soil edaphic properties, the microbial communities clustered according to the climatic condition of each study site (\cref{fig:figure_3b_label}). 
In semi-arid and mediterranean soils, microbial communities did not significantly change during the treatment. 
Both communities were similar and clustered apart from those of arid and humid soils along CCA 2. 
While microbial communities of the semi-arid and mediterranean soils were similar (along CCA 1), soil parameters of the respective sites differed significantly (along PC 1). 
In contrast, microbial communities in humid soils displayed an increasing variance along the y-axis (explaining \SI{24.3}{\percent} of the variance) with increasing number of W-D cycles and under constant moisture. 
Similarly, the microbial communities of R6 and CM in arid soils varied from R0--R3 along the x-axis (explaining \SI{26.5}{\percent}).

The microbial community composition was assessed by MiSeq amplicon sequencing. 
After quality check and filtering (including removing reads of the negative control) the dataset comprised 1.2 million reads across \(\sim\)\num{11000} amplicon sequencing variants (ASVs). 
From this \(\sim\)\SI{98.1}{\percent} of the reads were assigned to bacteria and \SI{1.9}{\percent} to archaea. 
After taxonomic classification, 828 putative genera (807 for bacteria, 21 for archaea) were obtained. 
Based on phylogenetic classification, ASVs could be assigned to the ten dominant prokaryotic phyla Proteobacteria (\SI{24.8}{\percent} of total ASVs), Actinobacteriota (\SI{22.9}{\percent}), Acidobacteriota (\SI{11.9}{\percent}), Chloroflexi (\SI{8.7}{\percent}), Planctomycetota (\SI{8.5}{\percent}), Bacteroidota (\SI{5.1}{\percent}), Gemmatimonadota (\SI{4.8}{\percent}), Verrucomicrobiota (\SI{3.6}{\percent}), Firmicutes (\SI{2.3}{\percent}) and Crenarchaeota (\SI{1.8}{\percent}) (\cref{fig:figure_a4_label}).

The bacterial communities of all four sites comprised 63 families and 18 unknown orders (\cref{fig:figure_6_label}, \cref{tab:table_a2_label}). 
Of these families, 32 possessed almost or more than \SI{2}{\percent} mean relative abundance per site. 
Some were abundant (\SI{> 2}{\percent}) only at one site, while others were found in soils from several climatic conditions. 
Euzebyaceae, Pseudonocardiaceae, Caldilineaceae, AKYG1722 and JG30-KF-CM45 (Thermomicrobiales), Trueperaceae, Longimicrobiaceae, Pirellulaceae and unclassified Rhizobiales were abundant exclusively in arid soils from Pan de Az{\'u}car. 
Nitrososphaeraceae, Geodermatophilaceae, Micromonosporaceae and Beijerinckiaceae were abundant in the semi-arid soils of Santa Gracia. 
Vicinamibacteraceae and unclassified Vicinamibacterales were found to be abundant in mediterranean soils of La Campana. 
In contrast, Subgroup 1 of Acidobacteriales, Acidothermaceae, Ktedonobacteraceae, Gemmataceae and unclassified Elsterales were abundant only in the humid soils of Nahuelbuta.

@@

Almost \num{11000} ASVs were assigned to 45 ecologically relevant functions using FAPROTAX, with 24 functions of significant abundance (cumulative abundance \SI{> 1}{\percent} across sites). 
In general, those predicted functions were related to organic matter degradation (chemoheterotrophy, fermentation, cellulolysis) phototrophy (anoxygenic photoautotrophy, photoautotrophy, photoheterotrophy and phototrophy), the nitrogen cycle (denitrification, nitrate denitrification, nitrate reduction, nitrate respiration, nitrification, nitrite denitrification, nitrite respiration, nitrogen fixation, nitrogen respiration, nitrous oxide denitrification) and degradation processes such as aromatic compound and hydrocarbon degradation as well as ureolysis. 
The most abundant predicted functions were related to chemoheterotrophy and aerobic chemoheterotrophy (\cref{fig:figure_a5_label}). 
In arid soils, additionally nitrate reduction (\SIrange{4}{7.5}{\percent}), nitrification and aerobic ammonia oxidation (\SIrange{0.5}{3.1}{\percent}) were enhanced (\cref{fig:figure_a5b_label}). 
Some predicted functions decreased after the first W-D cycle (aerobic heterotrophy, chemoheterotrophy, manganese oxidation, nitrate reduction, nitrate respiration and nitrogen respiration), while others were enhanced after 1 and/or 3 W-D cycles (phototrophy, aromatic compound degradation, cellulolysis, ureolysis and many nitrogen cycle-related functions such as denitrification, hydrocarbon degradation, denitrification, nitrite respiration, nitrogen fixation, nitrous oxide denitrification) (\cref{fig:figure_a5a_label}). 
In semi-arid soils, again nitrate reduction (\SIrange{2.4}{3.5}{\percent}), nitrification (\SIrange{3.1}{4.6}{\percent}) and aerobic ammonia oxidation (\SIrange{3.0}{5.8}{\percent}) were enhanced following heterotrophy (\cref{fig:figure_a5b_label}). 
Despite low abundances, many predicted functions were increased after 3 W-D cycles and most of them also in constantly moist soils. 
Similar observations were made in mediterranean soils. 
Humid soils were characterized by an enhanced abundance of ureolysis with an increasing trend, and cellulolysis as well as nitrogen fixation, with a decreasing trend during W-D treatment (\cref{fig:figure_a5b_label}). 
The decreasing abundance was observed also for several other functions, while aerobic ammonia oxidation, fermentation, nitrification, phototrophy and photoheterotrophy were most abundant under constant moisture conditions (\cref{fig:figure_a5a_label}).

@@

The four study sites along the Chilean Coastal Cordillera comprised soils ranging from poorly developed soils such as Regosol in the arid north to more developed soils such as Cambisols (semi-arid and mediterranean study sites) and to Umbrisols in the humid south. 
The soils were characterized by increasing clay, soil organic matter, soil moisture contents and soil acidity from north to south \citep{Bernhard2018b} (\cref{fig:figure_2_label}) and, thus provide different initial conditions for studying the microbial impact on soil aggregation and stabilization.

The level of soil development was not only indicated by the distribution of soil aggregate size classes but also by organic C and N contents of the particular soil aggregate classes. 
Overall, lowest SOC and TN contents were observed in the MAC fraction. 
They were higher in MIC and highest in the SC fraction. 
Clay and silt are known to build strong relations with stable humic substances \citep{Bouajila2021} and protect them against decomposition by microbial activity. 
Across sites, arid soils were characterized by maximum C/N ratios, typical for soil systems with a high amount of recalcitrant organic matter that is decomposed very slowly. 
Although, MIC are characterized by having increased amounts of biochemically recalcitrant C (i.e. compounds with a higher C/N ratio) \citep{Elliott1986, Kravchenko2015}, we observed highest C/N ratios in the MAC fractions of all four soil types. 
In accordance with previous work \citep{Kuramae2012}, the C/N ratio appeared to be a strong driver of the microbial community composition (\cref{fig:figure_3b_label}) as well as of abundance and diversity (\cref{fig:figure_a2_label} and \cref{fig:figure_a3_label}).

@@

Contrary to the arid soil material, microbial abundance as a measure of activity was not as strongly stimulated but remained relatively stable in semi-arid and mediterranean soil material during the manipulation experiment (\cref{fig:figure_5a_label}). 
Similar to the abundance, the prokaryotic diversity and community composition (\cref{fig:figure_3b_label}, \cref{fig:figure_6_label}) of both soil types were comparable. 
A main explanation could be the similar soil acidity \citep{Bernhard2018b}, which is controlled by the water balance as a result of the climate conditions \citep{Slessarev2016}. 
In addition, W-D of both soil types caused no significant changes of the community composition (\(p = 0.8\) in semi-arid, \(p = 1\) in mediterranean soil material). 
Despite the consistent community composition in both soil types, W-D resulted in a microbial response that had similar effects on the soil parameters (enhanced MIC and TN\textsubscript{SC} contents, but a loss of MAC), but different effects on aggregate stability.

>>>>

\begin{center}
  \textbf{\Large Manuscript 4}
\end{center}

@@

\begin{center}
  \textbf{\huge Microbially induced soil aggregate turnover across different climates and moisture regimes}
\end{center}

\begin{center}
  \textit{Geoderma}\\
  \textit{DOI 10.2139/ssrn.4740477}
\end{center}

\begin{justify}
    Julia Mitzscherling\textsuperscript{1}, Rómulo Oses\textsuperscript{2}, Nicolás Riveras-Muñoz\textsuperscript{3}, Carsten W. Müller\textsuperscript{4,5}, Oscar Seguel\textsuperscript{6},   Peter Kühn\textsuperscript{3}, Thomas Scholten\textsuperscript{3} and Dirk Wagner\textsuperscript{1,7}
\end{justify}

@@

  \vspace{0.4cm}

  \begin{flushleft}
    Received: 27 February 2024
  \end{flushleft}
  \cleardoublepage

@@

In the frame of the EarthShape project (German Science Foundation SPP 1803), soils were studied at the following four study sites distributed over \SI{1300}{\kilo\metre} along the Chilean Coastal Cordillera: Pan de Az{\'u}car National Park (\(\sim\)\ang{26}S), Santa Gracia Nature Reserve (\(\sim\)\ang{30}S), La Campana National Park (\(\sim\)\ang{33}S), and Nahuelbuta National Park (\(\sim\)\ang{38}S). The study sites cover a climate gradient with decreasing annual temperature and increasing annual precipitation from north to south and corresponded to arid, semi-arid, mediterranean and humid climates (\ref{fig:M4-F1}) resulting in different soil moisture regimes \citep{Bernhard2018}. Due to the long-term impact of stable climate conditions \citep{Ewing2006} soils show different stages of soil development. They were classified into Regosols in Pan de Az{\'u}car, Cambisols in Santa Gracia and La Campana, and Umbrisols in Nahuelbuta \citep{Bernhard2018}. Vegetation types range from very sparse higher plant desert vegetation in the north, where biocrust soil coverage can reach as much as \SI{40}{\percent} of the area (biocrust-dominated desert vegetation) over semi-desert open shrubland and mediterranean dry forest and to temperate mixed broadleaved-coniferous forest in the south \citep{Bernhard2018a, Bernhard2018}.

Surface depth increments (\SIrange{0}{5}{\centi\metre}) were collected in March--April 2016 on the mid slope position of the South-facing hillslopes in each study site and were directly sieved (\SI{< 2}{\milli\metre}) on ethanol-sterilized sieves to remove coarse organic material and stones. Samples were stored at \SI{4}{\degreeCelsius} until further processing. 
Detailed site-specific descriptions and soil classifications were performed by \citet{Bernhard2018a}, \citet{Bernhard2018} and \citet{Oeser2018a}, \citet{Oeser2018b}.

@@

Soil-chemical properties were analyzed for the native and the sterilized soils. The total organic carbon (TOC) was determined by Potsdamer Wasser- und Umweltlabor GmbH (PWU, Potsdam, Germany) according to DIN EN ISO/IEC 17025:2005. The aggregate size separation was done by wet sieving adapted from \citet{Elliott1986} as described in \citet{RiverasMunoz2022}. In the first step, \SI{10}{\gram} of soil material was irrigated with \SI{50}{\milli\liter} water. A series of four sieves, which were placed in descending order, was used to obtain the following soil aggregate size classes: (i) $>$\SI{500}{\micro\meter} (large macroaggregates); (ii) \SIrange{250}{500}{\micro\meter} (small macroaggregates); (iii) \SIrange{53}{250}{\micro\meter} (large microaggregates); (iv) \SIrange{20}{53}{\micro\meter} (small microaggregates), and (v) $<$\SI{20}{\micro\meter} (silt/clay).

Aggregates were separated manually by moving the sieve up and down 30 times during \SI{5}{\minute} in a \SI{2.5}{\liter} water tank. The aggregate classes retained on each sieve were oven-dried at \SI{105}{\degreeCelsius}, and weighed. In the following, large and small macroaggregates are referred to as macroaggregates (MAC), large and small microaggregates are referred to as microaggregates (MIC) and the silt/clay class is referred to as SC.

@@

Sequencing raw data are publicly available via the European Nucleotide Archive under the Project accession number \texttt{PRJEB72146}: \url{https://www.ebi.ac.uk/ena/browser/view/PRJEB72146}. All data used in this study are freely available under the Creative Commons Attribution 4.0 International (CC BY 4.0) open access license at GFZ data services. All samples are identified via a unique IGSN (international geo sample number). These data are available as a supplementary dataset in \citet{Bernhard2018} with the following DOI: \url{https://doi.org/10.5880/GFZ.5.3.2018.001}. When using the data please cite: \citet{Bernhard2018a}.

@@

Macroaggregates (MAC), microaggregates (MIC) and the silt/clay (SC) classes were differently distributed in arid, semi-arid, mediterranean and humid soils (Figure \ref{fig:M4-F2}a). Less developed soils from arid and semi-arid climate conditions were characterized by the lowest content of MAC ranging around \SIrange{60}{65}{\percent}. Semi-arid soils displayed a higher MIC content and a lower SC content compared to arid soils. More developed soils from the mediterranean and humid climates were characterized by the highest abundance of MAC with nearly \SI{80}{\percent} in mediterranean soils and more than \SI{90}{\percent} in humid soils. Accordingly, the proportion of MIC was low in mediterranean soils (\(\sim\)\SI{20}{\percent}) and lowest in humid soils (\SI{< 20}{\percent}). Moreover, SC was hardly present in mediterranean and humid soils (\(\sim\)\SI{2}{\percent}).

@@

The SOC and TN (Figure \ref{fig:M4-F2}b-c) contents were lowest in the less developed soils and increased with soil development along the climate gradient, with the lowest SOC (\(\sim\)\SI{1}{\percent}) and TN values (\(\sim\)\SI{0.1}{\percent}) in arid soils and highest values in humid soils (\(\sim\)\SI{10}{\percent} and \(\sim\)\SI{0.5}{\percent}). In each soil, the SOC and TN contents were lowest within the MAC class and highest within the SC class. The C/N ratios (Figure \ref{fig:M4-F2}d) were highest in arid soils. Lowest values were observed in semi-arid soils with an increasing trend over mediterranean to humid soils. Within each soil, the highest C/N ratios can be observed within MAC, followed by SC and MIC.

The aggregate stability (MWD) was lowest in arid and semi-arid soils, increased in mediterranean soils and peaked in humid soils.

A principal component analysis (PCA) including soil aggregate size classes, aggregate stability (MWD), SOC, TOC and TN content as well as the C/N determined during repeated W-D cycles (Figure \ref{fig:M4-F3}a) revealed similar soil properties in native and sterilized soils from the same climatic condition. The separation of soils along the x-axis, explaining \SI{66.0}{\percent} of the variance, was in accordance with the climate gradient from arid to humid conditions. The difference between arid and semi-arid soils was mainly visible on the y-axis, explaining \SI{27.6}{\percent} of the variance. In addition, native and sterilized soils from arid conditions were separated along the y-axis.

@@

Differences between native and sterilized soils were further evaluated in separate PCA ordinations, showing significant differences between native soils and sterile controls not only in arid (\(p = 0.005\)), but also in semi-arid (\(p = 0.008\)) and mediterranean soils (\(p = 0.02\)) (Figure \ref{fig:M4-F4}). At these three sites, the native samples with microorganisms before W-D treatment (\(\square\)R0) were similar to all sterile controls (\(\circ\)R0--R6, CM) and were thus referred to as microbially not (yet) influenced. After at least 1 W-D cycle, microbial samples clustered apart from the sterile controls and were thus claimed to be microbially influenced. The difference between microbially influenced and not influenced samples appeared mainly along the x-axis of the plots, which explained the highest variance. In arid soils, the microbially influenced soils were characterized by higher C/N ratios in all aggregate classes (Figure \ref{fig:M4-F4}a). In semi-arid and mediterranean soils, soil aggregate sizes in microbially influenced soils were characterized by higher levels of MIC and C/N\textsubscript{SC}, and lower levels of MAC compared to the sterile controls (Figure \ref{fig:M4-F4}b-c). Microbially influenced humid soils were not significantly different from sterile soils. However, the majority of microbially influenced soils was characterized by higher MAC and MWD, while sterile soils had higher MIC contents (Figure \ref{fig:M4-F4}d).

@@

Based on 16S and 28S rRNA gene copy numbers, bacteria were most abundant among all four climate conditions, followed by archaea and fungi (Figure \ref{fig:M4-F5}a). Only in humid soils, fungi were slightly more abundant than archaea and possessed highest abundances among soils. The lowest abundances of bacteria, archaea and fungi were observed in arid soils. Still, they increased over the course of the experiment with a peak after 3 cycles and maximum values under constant moisture. Semi-arid and mediterranean soils were characterized by highest abundances that remained relatively constant during the W-D cycles and constant moisture. In humid soils, bacterial and archaeal abundances were stable and fungal gene copies increased after 3 W-D cycles. However, all gene abundances of humid soils were lowest in soils exposed to constant moisture.

Linear (Pearson) and non-linear (Spearman) correlation analyses revealed that bacterial and archaeal abundances were positively correlated with each other. Both were inversely correlated with the C/N ratios in each soil class. On the contrary, the fungal abundance was positively correlated with MAC and MWD, while being inversely correlated with MIC and the S/C class. Furthermore, fungi were positively correlated with the SOC and TN contents in each aggregate class.

@@

Similar to their abundance, bacterial and archaeal diversity based on the alpha-diversity indices Chao-1 and Shannon (\(H\)) was highest in semi-arid soils from Santa Gracia (median \(H = 7.4\)) and in mediterranean soils of La Campana (\(H = 7.5\)) with maximum diversity indices in mediterranean soils (Figure \ref{fig:M4-F5}b). Although lowest microbial abundances were observed in arid soils, the bacterial and archaeal diversity was lowest in humid soils from Nahuelbuta (\(H = 6.0\)), followed by arid soils from Pan de Az{\'u}car (\(H = 6.6\)). In arid and semi-arid soils, the diversity according to Chao-1 increased after one W-D cycle and decreased afterwards. Contrary, mediterranean and humid soils were characterized by an increasing diversity during W-D cycles, but possessed a low diversity in the sample exposed to constant moisture. Like the bacterial and archaeal abundance, the number of taxa and the diversity were negatively correlated with the C/N ratios in all soil classes.

@@

Like the soil edaphic properties, the microbial communities clustered according to the climatic condition of each study site (Figure \ref{fig:M4-F3}b). In semi-arid and mediterranean soils, microbial communities did not significantly change during the treatment. Both communities were similar and clustered apart from those of arid and humid soils along CCA 2. While microbial communities of the semi-arid and mediterranean soils were similar (along CCA 1), soil parameters of the respective sites differed significantly (along PC 1). In contrast, microbial communities in humid soils displayed an increasing variance along the y-axis (explaining \SI{24.3}{\percent} of the variance) with increasing number of W-D cycles and under constant moisture. Similarly, the microbial communities of R6 and CM in arid soils varied from R0--R3 along the x-axis (explaining \SI{26.5}{\percent}).

The microbial community composition was assessed by MiSeq amplicon sequencing. After quality check and filtering (including removing reads of the negative control) the dataset comprised 1.2 million reads across \(\sim\)\num{11000} amplicon sequencing variants (ASVs). From this \(\sim\)\SI{98.1}{\percent} of the reads were assigned to bacteria and \SI{1.9}{\percent} to archaea. After taxonomic classification, 828 putative genera (807 for bacteria, 21 for archaea) were obtained. Based on phylogenetic classification, ASVs could be assigned to the ten dominant prokaryotic phyla Proteobacteria (\SI{24.8}{\percent} of total ASVs), Actinobacteriota (\SI{22.9}{\percent}), Acidobacteriota (\SI{11.9}{\percent}), Chloroflexi (\SI{8.7}{\percent}), Planctomycetota (\SI{8.5}{\percent}), Bacteroidota (\SI{5.1}{\percent}), Gemmatimonadota (\SI{4.8}{\percent}), Verrucomicrobiota (\SI{3.6}{\percent}), Firmicutes (\SI{2.3}{\percent}) and Crenarchaeota (\SI{1.8}{\percent}).

The bacterial communities of all four sites comprised 63 families and 18 unknown orders (Figure \ref{fig:M4-F6}). 
Of these families, 32 possessed almost or more than \SI{2}{\percent} mean relative abundance per site. 
Some were abundant (\SI{> 2}{\percent}) only at one site, while others were found in soils from several climatic conditions. 
Euzebyaceae, Pseudonocardiaceae, Caldilineaceae, AKYG1722 and JG30-KF-CM45 (Thermomicrobiales), Trueperaceae, Longimicrobiaceae, Pirellulaceae and unclassified Rhizobiales were abundant exclusively in arid soils from Pan de Az{\'u}car. 
Nitrososphaeraceae, Geodermatophilaceae, Micromonosporaceae and Beijerinckiaceae were abundant in the semi-arid soils of Santa Gracia. 
Vicinamibacteraceae and unclassified Vicinamibacterales were found to be abundant in mediterranean soils of La Campana. 
In contrast, Subgroup 1 of Acidobacteriales, Acidothermaceae, Ktedonobacteraceae, Gemmataceae and unclassified Elsterales were abundant only in the humid soils of Nahuelbuta.

@@

Almost \num{11000} ASVs were assigned to 45 ecologically relevant functions using FAPROTAX, with 24 functions of significant abundance (cumulative abundance \SI{> 1}{\percent} across sites). 
In general, those predicted functions were related to organic matter degradation (chemoheterotrophy, fermentation, cellulolysis) phototrophy (anoxygenic photoautotrophy, photoautotrophy, photoheterotrophy and phototrophy), the nitrogen cycle (denitrification, nitrate denitrification, nitrate reduction, nitrate respiration, nitrification, nitrite denitrification, nitrite respiration, nitrogen fixation, nitrogen respiration, nitrous oxide denitrification) and degradation processes such as aromatic compound and hydrocarbon degradation as well as ureolysis. 
The most abundant predicted functions were related to chemoheterotrophy and aerobic chemoheterotrophy. In arid soils, additionally nitrate reduction (\SIrange{4}{7.5}{\percent}), nitrification and aerobic ammonia oxidation (\SIrange{0.5}{3.1}{\percent}) were enhanced. Some predicted functions decreased after the first W-D cycle (aerobic heterotrophy, chemoheterotrophy, manganese oxidation, nitrate reduction, nitrate respiration and nitrogen respiration), while others were enhanced after 1 and/or 3 W-D cycles (phototrophy, aromatic compound degradation, cellulolysis, ureolysis and many nitrogen cycle-related functions such as denitrification, hydrocarbon degradation, denitrification, nitrite respiration, nitrogen fixation, nitrous oxide denitrification). In semi-arid soils, again nitrate reduction (\SIrange{2.4}{3.5}{\percent}), nitrification (\SIrange{3.1}{4.6}{\percent}) and aerobic ammonia oxidation (\SIrange{3.0}{5.8}{\percent}) were enhanced following heterotrophy. Despite low abundances, many predicted functions were increased after 3 W-D cycles and most of them also in constantly moist soils. Similar observations were made in mediterranean soils. Humid soils were characterized by an enhanced abundance of ureolysis with an increasing trend, and cellulolysis as well as nitrogen fixation, with a decreasing trend during W-D treatment. The decreasing abundance was observed also for several other functions, while aerobic ammonia oxidation, fermentation, nitrification, phototrophy and photoheterotrophy were most abundant under constant moisture conditions.

@@

The four study sites along the Chilean Coastal Cordillera comprised soils ranging from poorly developed soils such as Regosol in the arid north to more developed soils such as Cambisols (semi-arid and mediterranean study sites) and to Umbrisols in the humid south. The soils were characterized by increasing clay, soil organic matter, soil moisture contents and soil acidity from north to south \citep{Bernhard2018} (\ref{fig:M4-F2}) and, thus provide different initial conditions for studying the microbial impact on soil aggregation and stabilization.

The level of soil development was not only indicated by the distribution of soil aggregate size classes but also by organic C and N contents of the particular soil aggregate classes. Overall, lowest SOC and TN contents were observed in the MAC fraction. They were higher in MIC and highest in the SC fraction. Clay and silt are known to build strong relations with stable humic substances \citep{Bouajila2021} and protect them against decomposition by microbial activity. Across sites, arid soils were characterized by maximum C/N ratios, typical for soil systems with a high amount of recalcitrant organic matter that is decomposed very slowly. Although, MIC are characterized by having increased amounts of biochemically recalcitrant C (i.e. compounds with a higher C/N ratio) \citep{Elliott1986, Kravchenko2015}, we observed highest C/N ratios in the MAC fractions of all four soil types. In accordance with previous work \citep{Kuramae2012}, the C/N ratio appeared to be a strong driver of the microbial community composition (Figure \ref{fig:M4-F3}b) as well as of abundance and diversity.

@@

Contrary to the arid soil material, microbial abundance as a measure of activity was not as strongly stimulated but remained relatively stable in semi-arid and mediterranean soil material during the manipulation experiment (Figure \ref{fig:M4-F5}a). Similar to the abundance, the prokaryotic diversity and community composition (Figure \ref{fig:M4-F3}b, Figure \ref{fig:M4-F6}) of both soil types were comparable. A main explanation could be the similar soil acidity \citep{Bernhard2018}, which is controlled by the water balance as a result of the climate conditions \citep{Slessarev2016}. In addition, W-D of both soil types caused no significant changes of the community composition (\(p = 0.8\) in semi-arid, \(p = 1\) in mediterranean soil material). Despite the consistent community composition in both soil types, W-D resulted in a microbial response that had similar effects on the soil parameters (enhanced MIC and TN\textsubscript{SC} contents, but a loss of MAC), but different effects on aggregate stability.
